---
title: 'Stats 141SL '
author: "Rose Din (ID: 804764627), Lisa Kaunitz (ID: 404731988), Claire Markey (ID: 104789928), Allison Ocelo (ID: 404799327), Cristina Sanchez (ID: 805119924), and Kayla Schroeder (ID: 804729436) - Section 1A"
header-includes:
- \usepackage[usenames ,dvipsnames ,svgnames ,table]{xcolor}
- \usepackage{fontspec}
- \newfontfamily\comic[Scale=2]{ComicSansMS}
- \usepackage{graphicx}
- \usepackage{tabularx}
- \usepackage{enumitem}
- \usepackage{amsmath,amssymb,amsthm}
date: "03/05/2020"
output: 
  pdf_document:
    latex_engine: xelatex
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\thispagestyle{empty}
\newpage
\section{Abstract}

\textit{This is my abstract.}

\newpage
\section{Research Questions}

1. What are the significant risk factors that predispose a child to asd behaviors?

2. Is there a significant difference when looking at predictors for autism in males vs females? 

\section{Variables of the Study and how they were Measured}

\begin{center}
\begin{tabular}{ | l | l | p{10cm} |} 
\hline
\multicolumn{3}{|c|}{Data Codebook} \\
\hline
Number & Attribute & Description \\
\hline
\hline
1 & momsdeliveryage & Mother's Age at Delivery.\\ \hline
2 & momsedu & Mother's Education Level.	\\ \hline
3 & delivery & How was your baby delivered? \\ \hline
4 & babymortality & Misscarriages, or infant death at birth \\ \hline
5 & mathealth count & Maternal mental health count\\ \hline
6 & sex & Baby's Gender. 1 = Male, 0 = Female	\\ \hline
7 & tri alcohol & Total number of trimesters alcohol was used.	\\ \hline
8 & tri tabacco & Total number of trimesters tabacco was used. \\ \hline	
9 & tri vaping & Total number of trimesters vaping was used.\\ \hline
10 & tri cannabis & Total number of trimesters cannabis was used.\\ \hline		   
11 & mom trauma & Did the mother experience any extremely stressful event during pregnancy, such as the death of a family member, divorce, homelessness, living in a war zone, or abuse?\\ \hline	
12 & asd1 & Total number of family members with ASD (experiencing other)	
1 = have 1+ family members with ASD, 0 otherwise.\\ \hline		   
13 & szbpfamily1 & Total number of family members with SZ/BP (experiencing other). 
1 = have 1+ family members with SZ/BP, 0 otherwise. \\ \hline	
14 & ddfamily1 &Did any of the following family members have a history of developmental delay while an infant?
1 = have 1+ family members with developmental delay, 0 otherwise\\ \hline	
15 & artmethod 1  & In vitro fertilization (IVF) assisted in conceiving baby.	\\ \hline
16 & whenfever 3 & Mothers fever during Third Trimester (weeks 28-40). \\ \hline
29 & deaf1 & Total number of family members Deaf (experiencing other).
1 = have 1+ family members Deaf, 0 otherwise.\\ \hline	
17 & maternalpregnancyproblems 1 & Were there any of the following interventions, complications, or abnormalities during this pregnancy with baby - Gestational diabetes. \\ \hline
18 & maternalpregnancyproblems 2 & Were there any of the following interventions, complications, or abnormalities during this pregnancy with baby - Infections requiring antibiotics . \\ \hline
19 & maternalpregnancyproblems 3 & Were there any of the following interventions, complications, or abnormalities during this pregnancy with baby - Placenta previa. \\ \hline
20 & maternalpregnancyproblems 4 & Were there any of the following interventions, complications, or abnormalities during this pregnancy with baby - Antidepressants to treat depression. \\ \hline
21 & maternalpregnancyproblems 7 & Were there any of the following interventions, complications, or abnormalities during this pregnancy with baby - Preeclampsia/eclampsia . \\ \hline
22 & maternalpregnancyproblems 8 & Were there any of the following interventions, complications, or abnormalities during this pregnancy with baby -  Anemia. \\ \hline
23 & maternalpregnancyproblems 9 & Were there any of the following interventions, complications, or abnormalities during this pregnancy with baby - Hypertension (high blood pressure). \\ \hline
24 & maternalpregnancyproblems 11 & Were there any of the following interventions, complications, or abnormalities during this pregnancy with baby - Preterm labor. \\ \hline
25 & maternalpregnancyproblems 13 & Were there any of the following interventions, complications, or abnormalities during this pregnancy with baby - Placental abruption. \\ \hline
26 & maternalpregnancyproblems 14 & Were there any of the following interventions, complications, or abnormalities during this pregnancy with baby - Vaginal bleeding. \\ \hline
27 & maternalpregnancyproblems 16 & Were there any of the following interventions, complications, or abnormalities during this pregnancy with baby -  Hyperemesis gravidarum. \\ \hline
28 & risk score ASD & Total number of ASD risk behaviors.  Large number suggests higher risk.\\ 
\hline 
\end{tabular} 
\end{center}


\newpage
\section{Exploratory Data Analysis}

## Controlling for Sex: 

### Male 

### Female



\newpage
\section{Statistical analysis used to answer the research questions}
A random forest model was run on prenatal factors to determine which variables were important risk factors for `risk_score_ASD`. Highly correlated variables were dropped from the data before running the random forest model. Important variables were determined from the node purity and mean squared error measures, 14 total variables were selected as important risk factors from `risk_score_ASD`. These variables were included in the final Poisson model. A Poisson Model was chosen since we are using count data and the dispersion parameter was less than 1, in other words our variance was less than our expected value for all of our models.

```{r,echo=FALSE}
library(readr)
predata <- read_csv("~/predata.csv")

library(randomForest)

drops <- c("maternalpregnancyproblems___19", "maternalpregnancyproblems___20", "matmentalhealth___1", "matmentalhealth___2", "matmentalhealth___3","matmentalhealth___4", "matmentalhealth___5", "matmentalhealth___6", "matmentalhealth___7", "mat_dep", "mat_anx", "mat_ocd","mat_bp","mat_adhd","mat_pssd","maternalproblems1","art","momlivebirths","totalpregnancies","maternalpregnancyproblems___18","momfever")
predata1 <- na.omit(predata[ , !(names(predata) %in% drops)])

rf.pre <- randomForest(risk_score_ASD~.,data=predata1,mtry=16,importance=TRUE,ntrees=1000)
```
```{r,echo=FALSE}
library(randomForestExplainer)
min_depth_frame <- min_depth_distribution(rf.pre)
plot_min_depth_distribution(min_depth_frame, mean_sample = "relevant_trees", k = 15)
```
\textit{Text}


\newpage
\section{Summary of results}

\textit{Text}


\newpage
\section{Interpretation of results}

\textit{Text}

\newpage
\section{Interpretation of plots}

\textit{Text}

\newpage
\section{Overall conclusions}

\textit{Text}

\newpage
\section{Challenges of the study}

\textit{Text}

\newpage
\section{Recommendations for the future}

\textit{Text}
